sWGS validation report

### 2019-08-14 Insilico Splits.Rmd
...

cache.dir = paste(data, 'Analysis/5e6_arms_all', sep='/')
split.dir = paste(data, 'Analysis/5e6_arms_splits', sep='/')

@@

roc = pROC::roc(Status ~ Prediction, data=predictions, auc=T, ci=T, of='thresholds')
roc$model = 'discovery'
roc$is = 'discovery'

@@

mrplot <- function(rocList) {
  aucs = do.call(rbind, lapply(rocList,function(r) 
    cbind.data.frame('AUC'=r$auc*100,'model'=r$model,'Specificity'=pROC::coords(r,'best')[['specificity']], 'Sensitivity'=pROC::coords(r,'best')[['sensitivity']])))
  aucs$y = aucs$Sensitivity
  aucs$x = aucs$Specificity

  df = do.call(rbind, lapply(rocList, function(r) 
    cbind.data.frame('Specificity'=rev(r$specificities), 'Sensitivity'=rev(r$sensitivities),'model'=r$model, 'is'=r$is) 
  ))

  ggplot(df, aes(Specificity, Sensitivity,color=is)) + geom_line(size=1) + 
    geom_line(data=subset(df, model == 'discovery'), color='red', size=1) +
    scale_x_reverse() +
    scale_color_brewer(palette='Blues') +
    labs(title='', x='Specificity (FPR)', y='Sensitivity (TPR)')  + plot.theme +
    theme(legend.position = 'none')
}

>>>>

cache.dir = paste(data, 'Analysis/5e6_arms', sep='/')
split.dir = paste(data, 'Analysis/5e6_arms_splits', sep='/')

@@

predictions = pg.samp

roc = pROC::roc(Status ~ Prediction, data=predictions, auc=T, ci=T, of='thresholds', transpose = T)
roc$model = 'discovery'
roc$is = 'discovery'

@@

mrplot <- function(rocList) {
  aucs = do.call(bind_rows, lapply(names(rocList),function(nm) { 
    r = rocList[[nm]]
    tibble( AUC = r$auc*100, model = r$model, 
      'Specificity'= pROC::coords(r,'best', ret='specificity',transpose=T)[1],
      'Sensitivity'= pROC::coords(r,'best', ret='sensitivity',transpose=T)[1] )
  }))

  aucs$y = aucs$Sensitivity
  aucs$x = aucs$Specificity

  df = do.call(bind_rows, lapply(names(rocList),function(nm) { 
  #  print(nm)
    r = rocList[[nm]]
    tibble('Specificity'=rev(r$specificities), 'Sensitivity'=rev(r$sensitivities),'model'=r$model, 'is'=as.character(r$is) )
  }))

  ggplot(df, aes(Specificity, Sensitivity,color=is)) + geom_line(color='lightblue') + 
    geom_line(data=subset(df, model == 'discovery'), color='red', size=1) +
    scale_x_reverse() +
    #scale_color_brewer(palette='Blues') +
    labs(title='', x='Specificity (FPR)', y='Sensitivity (TPR)')  + plot.theme +
    theme(legend.position = 'none')
}

### 2019-08-14 Paper.Rmd
...

patient.info = read.patient.info(patient.file, demo.file, set='All')$info
head(patient.info)

@@

patient.info = arrange(patient.info, Status, Hospital.Research.ID, Endoscopy.Year, Pathology)
sum.patient.data = summarise.patient.info(patient.info)

@@

```{r labelsPNP, echo=F, message=F, include=T}
cache.dir = paste(data, 'Analysis/5e6_arms_all', sep='/') # QC'd

@@

chr.lengths = get.chr.lengths()
chr.lengths$chrom = sub('chr','',chr.lengths$chrom)
chr.lengths = subset(chr.lengths, chrom %in% c(1:22))
chr.lengths$chrom = factor(chr.lengths$chrom, levels=c(1:22), ordered = T)

@@

```{r back, echo=F, message=F, warning=F}
transform.by.time<-function(pt, info) {
  #pt = left_join(pt, info %>% dplyr::select(Hospital.Research.ID, Patient,Samplename,Block), by=c('Hospital.Research.ID','Patient','Samplename'))

  pt = pt %>% dplyr::group_by(PID, Samplename) %>% dplyr::mutate( 'months.before.final' = (Final.Endoscopy - Endoscopy.Year)*12) %>% arrange(Pathology, PID, Endoscopy.Year) %>% ungroup

@@

back$quants = with(back, cut(Prediction, breaks=cuts))
qt = as.data.frame.matrix(table(back$quants, back$Status))
qt$quant = rownames(qt)

@@

pcfit = lm(perc~mn, data=pred.confidence)
rsid = rstandard(pcfit)
#pred.confidence[which( rsid >= 1 |  rsid <= -1 ), 'Risk'] = 'Moderate'
pred.confidence = pred.confidence %>% mutate(Risk = as.character(ifelse(p.value > 0.05, 'Moderate', Risk)))

@@

```{r pcalib, echo=F, warning=F, message=F, fig.width=8, fig.height=6}
pcal = BarrettsProgressionRisk::showPredictionCalibration(pred.confidence)
pcal
ggsave(filename='plots/pred_calib.png', plot=pcal, width=4, height=4, units='in', dpi=300)
```

@@

list.files(cache.dir)

@@

@@

`r pander(as.data.frame(summary.t))`

@@

```{r val_setup, echo=F, message=F, warning=F, include=F}
validation.patient.info =  subset(all.patient.info, Set == 'Test')
validation.patient.info = arrange(validation.patient.info, Status, Hospital.Research.ID, Endoscopy.Year, Pathology)

@@

cleaned = list.files(path=paste(data, 'Cleaned', sep='/'), pattern='tiled', full.names=T, recursive=T)
cleaned = grep(paste(unique(validation.patient.info$Hospital.Research.ID), collapse='|'), cleaned, value=T)

raw = list.files(path=paste(data, 'Cleaned', sep='/'), pattern='raw', full.names=T, recursive=T)
raw = grep(paste(unique(validation.patient.info$Hospital.Research.ID), collapse='|'), raw, value=T)

# Get the SD, mean, and variance(?) per sample
v.sampleVar = do.call(rbind, lapply(raw, function(f) {
  t = read.table(f, sep='\t', header=T)
  sampleCols = grep('chr|arm|start|end|probes', colnames(t), invert = T)
  x = do.call(rbind, lapply(sampleCols, function(i){
    cbind('Sample.Mean'=mean(t[,i]),
          'Sample.SD'=sd(t[,i]),
          'Sample.Var'=var(t[,i]))
  }))
  rownames(x) = colnames(t)[sampleCols]
  return(x)
}))

arms = grep('arms', cleaned, value=T)
segs = grep('arms', cleaned, invert=T, value=T)

length(segs) == length(arms)

# Load segment files
tiled.segs = do.call(rbind, lapply(segs, function(f) {
  load.segment.matrix(f)
}))
dim(tiled.segs)

tiled.arms = do.call(rbind, lapply(arms, function(f) {
  load.segment.matrix(f)
}))

length(z.mean) == ncol(tiled.segs)

for (i in 1:ncol(tiled.segs)) 
  tiled.segs[,i] = unit.var(tiled.segs[,i], z.mean[i], z.sd[i])

for (i in 1:ncol(tiled.arms)) 
  tiled.arms[,i] = unit.var(tiled.arms[,i], z.arms.mean[i], z.arms.sd[i])

dim(tiled.arms)
dim(tiled.segs)

cx.score = unit.var(score.cx(tiled.segs,1), mn.cx, sd.cx)

validation.df = subtract.arms(tiled.segs, tiled.arms)
validation.df = cbind(validation.df, 'cx'=cx.score)

fauxlabel = rep(0, nrow(validation.df) )
names(fauxlabel) = intersect(validation.patient.info$Samplename, rownames(validation.df))

@@

```{r validation_globalplot, message=F, warning=F, echo=F, fig.height=8, fig.width=8}
valCNdf = t(validation.df[,grep('^\\d+:',colnames(validation.df))])

@@

```{r validation, message=F, warning=F, echo=F, fig.height=8, fig.width=8}
file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
message(paste("loading", file))
load(file, verbose=T)

fitV = models[[select.alpha]]
lambda.opt = performance.at.1se[[select.alpha]][, 'lambda']

pred = predict(fitV, newx=validation.df, s=lambda.opt, type='response')
or = predict(fitV, newx=validation.df, s=lambda.opt, type='link')

vpd = lapply(unique(validation.patient.info$Hospital.Research.ID), function(pt) {
  info = subset(validation.patient.info, Hospital.Research.ID == pt)
  info$Prediction = NA
  info$OR = NA
  info$PID = unlist(lapply(info$Path.ID, function(x) unlist(strsplit(x, 'B'))[1]))

  info = cbind(info, v.sampleVar[info$Samplename,])

  return(info)
})
names(vpd) = unique(validation.patient.info$Hospital.Research.ID)

@@

for (pt in names(vpd)) {
  vpd[[pt]]$Prediction[which(vpd[[pt]]$Samplename %in% rownames(pred))] = pred[ intersect(rownames(pred), vpd[[pt]]$Samplename), ]
  vpd[[pt]]$OR[which(vpd[[pt]]$Samplename %in% rownames(or))] = or[ intersect(rownames(or), vpd[[pt]]$Samplename), ]
}

back.v = do.call(rbind.data.frame,lapply(vpd, transform.by.time, info=validation.patient.info))
back.v = merge(back.v, validation.patient.info[,c('Samplename','p53.Status')])

sq = c(0, seq(1,132,12),max(back$months.before.final))
back.v$brks = cut(back.v$months.before.final, 
                  breaks=sq, 
                 labels=c(c('Endpoint',apply(cbind(sq[-1], sq[-1]+11), 1, function(x) paste(x, collapse='-')))[-c(length(sq)-1,length(sq))], '>120'),
                  include.lowest = T, ordered_result = T)
back.v$brks = with(back.v, factor(brks, rev(levels(brks))))

back.v$Pathology = as.character(back.v$Pathology)

back.v$Pathology = factor(back.v$Pathology, levels=c('BE','ID','LGD','HGD','IMC'), ordered = T)

@@

```{r echo=F, message=F, warning=F, fig.height=5, fig.width=8}
back.v$quants = with(back.v, cut(Prediction, breaks=cuts))
back.v = merge(back.v, pred.confidence[c('quant','perc','Risk')], by.x='quants', by.y='quant')
back.v$Risk = factor(back.v$Risk, levels=c('High','Moderate','Low'))
back.v$nl = factor(back.v$nl, ordered = T)

@@

back.v$Prediction

table(subset(back.v, Status == 'P')$Prediction > 0.6)
table(subset(back.v, Status == 'NP')$Prediction < 0.3)

@@

```{r, eval=F}
# TODO: Relative risk
pred = predict(f, newx=dysplasia.df, s=performance.at.1se$`0.9`$lambda)
?predict
?predict.glm
quantile(pred)
p0 = predict(f, newx=dysplasia.df, s=performance.at.1se$`0.9`$lambda, type='response')
p0
p1 = 1/(1+exp(-pred+2.66))
plot(p0,p1)
hist(p1)
hist(p0)
hist(p1)
plot(p0,p1)
hist(p1, n=20)
mean(p1>.5)
mean(p0>.5)
hist(pred)

>>>>

all.patient.info = read.patient.info(patient.file, demo.file, set='All')$info

patient.info = all.patient.info %>% dplyr::filter(Set == 'Training')
head(patient.info)

@@

patient.info = arrange(patient.info, Status, Hospital.Research.ID, Endoscopy.Year, Pathology)
sum.patient.data = as_tibble(summarise.patient.info(patient.info))

@@

```{r labelsPNP, echo=F, message=F, include=T}
cache.dir = paste(data, 'Analysis/5e6_arms', sep='/') # QC'd

@@

chr.lengths = BarrettsProgressionRisk:::chrInfo(prefix = '') %>% dplyr::filter(chr %in% c(1:22)) %>%
  dplyr::mutate(chrom = factor(sub('chr','', chrom), levels=c(1:22), ordered=T))
#chr.lengths$chrom = sub('chr','',chr.lengths$chrom)
#chr.lengths = subset(chr.lengths, chrom %in% c(1:22))
#chr.lengths$chrom = factor(chr.lengths$chrom, levels=c(1:22), ordered = T)

@@

```{r back, echo=F, message=F, warning=F}
transform.by.time<-function(pt, info) {
  #pt = left_join(pt, info %>% dplyr::select(Hospital.Research.ID, Patient,Samplename,Block), by=c('Hospital.Research.ID','Patient','Samplename'))
  pt = pt %>% dplyr::group_by(PID, Samplename) %>% dplyr::mutate( 'months.before.final' = (Final.Endoscopy - Endoscopy.Year)*12) %>% arrange(Pathology, PID, Endoscopy.Year) %>% ungroup

@@

back$quants = with(back, cut(Prediction, breaks=cuts, include.lowest = T))
qt = as.data.frame.matrix(table(back$quants, back$Status))
qt$quant = rownames(qt)

@@

pcfit = lm(perc~mn, data=pred.confidence)
rsid = rstandard(pcfit)
#pred.confidence[which( rsid >= 1 |  rsid <= -1 ), 'Risk'] = 'Moderate'
pred.confidence = pred.confidence %>% mutate(Risk = as.character(ifelse(p.value > 0.05 & r1 < 0.6, 'Moderate', Risk)))

@@

```{r pcalib, echo=F, warning=F, message=F, fig.width=8, fig.height=6}
pcal = BarrettsProgressionRisk::showPredictionCalibration(pred.confidence)
pcal
ggsave(filename='plots/pred_calib.png', plot=pcal, width=4, height=4, units='in', dpi=300)

back.per.endo = back %>% group_by(Hospital.Research.ID, Status, Patient, PID) %>% 
  dplyr::mutate('n.samples'=length(PID),'Samples'=paste(Samplename,collapse=',')) %>% 
  dplyr::summarise_at(vars(matches('Pred|RR|Year')), list(max)) %>% dplyr::select(-matches('^Sample$')) %>% 
  arrange(Patient, Endoscopy.Year, PID) 
back.per.endo = back.per.endo %>% mutate(quants = cut(Prediction, breaks=cuts, include.lowest = T)) %>% left_join(pred.confidence %>% dplyr::select(quants, Risk), by='quants') %>% dplyr::select(-matches('quant')) %>% dplyr::mutate(Risk = as.character(Risk))

back %>% group_by(Status, Risk)  %>% dplyr::summarise(n = length(Risk)) %>% spread(Risk, n) %>% rowwise %>% 
  dplyr::mutate(n=sum(High,Low,Moderate), Low = Low/n, Moderate = Moderate/n, High = High/n) %>%
  mutate_if(is.double, list(round),2) %>%
  kableExtra::kable(caption='Training set per-sample predictions') %>% kableExtra::kable_styling(full_width = F)

back.per.endo %>%  group_by(Status, Risk)  %>% dplyr::summarise(n = length(Risk)) %>% spread(Risk, n) %>% rowwise %>% 
  dplyr::mutate(n = sum(High,Low,Moderate), Low = Low/n, Moderate = Moderate/n, High = High/n) %>% mutate_if(is.double, list(round),2) %>%
  kableExtra::kable(caption='Training set per-endoscopy predictions') %>% kableExtra::kable_styling(full_width = F)
```

@@

#list.files(cache.dir)

@@

summary.t %>% kableExtra::kable() %>% kableExtra::kable_styling('striped', full_width = F)

@@

@@

```{r val_setup, echo=F, message=F, warning=F, include=F}
validation.patient.info =  all.patient.info %>% filter(Set == 'Test') %>% arrange(Status, Hospital.Research.ID, Endoscopy.Year, Pathology)

@@

predicted = '~/Data/BarrettsProgressionRisk/Analysis/training/val_predictions'

vpd = do.call(bind_rows, lapply(list.files(predicted, full.names = T), function(f) {
  read_tsv(f, col_types = 'cddcccdccdc')
})) %>% dplyr::select(Hospital.Research.ID, everything())

vpd = left_join(vpd %>% dplyr::select(-Pathology, -matches('P53')), validation.patient.info %>% dplyr::select(Samplename, Patient, Pathology, p53.Status, Block, Initial.Endoscopy, Final.Endoscopy, Status ), by=c('Sample'='Samplename')) %>% 
  dplyr::rename(Endoscopy.Year = Endoscopy, Samplename = Sample, RR = `Relative Risk`, Prediction = Probability)

vpd.per.endo = vpd %>% group_by(Hospital.Research.ID, Status, Patient, PID) %>% 
  dplyr::mutate('n.samples'=length(PID),'Samples'=paste(Samplename,collapse=',')) %>% 
  dplyr::summarise_at(vars(matches('Pred|RR|Year')), list(max)) %>% dplyr::select(-matches('^Sample$')) %>% 
  arrange(Patient, Endoscopy.Year, PID) 
vpd.per.endo = vpd.per.endo %>% mutate(quants = cut(Prediction, breaks=cuts, include.lowest = T)) %>% left_join(pred.confidence %>% dplyr::select(quants, Risk), by='quants') %>% dplyr::select(-matches('quant')) %>% dplyr::mutate(Risk = as.character(Risk))

vpd %>% group_by(Status, Risk)  %>% dplyr::summarise(n = length(Risk)) %>% spread(Risk, n) %>% rowwise %>% 
  dplyr::mutate(n=sum(High,Low,Moderate), Low = Low/n, Moderate = Moderate/n, High = High/n) %>%
  mutate_if(is.double, list(round),2) %>%
  kableExtra::kable(caption='Test set per-sample predictions') %>% kableExtra::kable_styling(full_width = F)

vpd.per.endo %>%  group_by(Status, Risk)  %>% dplyr::summarise(n = length(Risk)) %>% spread(Risk, n) %>% rowwise %>% 
  dplyr::mutate(n = sum(High,Low,Moderate), Low = Low/n, Moderate = Moderate/n, High = High/n) %>% mutate_if(is.double, list(round),2) %>%
  kableExtra::kable(caption='Test set per-endoscopy predictions') %>% kableExtra::kable_styling(full_width = F)

# Get the SD, mean, and variance(?) per sample
# v.sampleVar = do.call(rbind, lapply(raw, function(f) {
#   t = read.table(f, sep='\t', header=T)
#   sampleCols = grep('chr|arm|start|end|probes', colnames(t), invert = T)
#   x = do.call(rbind, lapply(sampleCols, function(i){
#     cbind('Sample.Mean'=mean(t[,i]),
#           'Sample.SD'=sd(t[,i]),
#           'Sample.Var'=var(t[,i]))
#   }))
#   rownames(x) = colnames(t)[sampleCols]
#   return(x)
# }))

# # Load segment files
# tiled.segs = do.call(rbind, lapply(segs, function(f) {
#   load.segment.matrix(f)
# }))
# dim(tiled.segs)
# 
# tiled.arms = do.call(rbind, lapply(arms, function(f) {
#   load.segment.matrix(f)
# }))
# 
# length(z.mean) == ncol(tiled.segs)
# 
# for (i in 1:ncol(tiled.segs)) 
#   tiled.segs[,i] = unit.var(tiled.segs[,i], z.mean[i], z.sd[i])
# 
# for (i in 1:ncol(tiled.arms)) 
#   tiled.arms[,i] = unit.var(tiled.arms[,i], z.arms.mean[i], z.arms.sd[i])
# 
# dim(tiled.arms)
# dim(tiled.segs)
# 
# cx.score = unit.var(score.cx(tiled.segs,1), mn.cx, sd.cx)
# 
# validation.df = subtract.arms(tiled.segs, tiled.arms)
# validation.df = cbind(validation.df, 'cx'=cx.score)
# 
# fauxlabel = rep(0, nrow(validation.df) )
# names(fauxlabel) = intersect(validation.patient.info$Samplename, rownames(validation.df))

@@

```{r validation_globalplot, message=F, warning=F, echo=F, fig.height=8, fig.width=8, eval=F}
valCNdf = t(validation.df[,grep('^\\d+:',colnames(validation.df))])

@@

```{r validation, message=F, warning=F, echo=F, fig.height=8, fig.width=8}

@@

back.v = do.call(bind_rows, lapply(unique(vpd$Patient) %>% map( function(p) vpd %>% filter(Patient == p) ), transform.by.time, info=validation.patient.info)) %>% dplyr::mutate(Patient = factor(Patient))

sq = c(0, seq(1,132,12),max(back$months.before.final))
back.v$brks = cut(back.v$months.before.final, 
                  breaks=sq, labels=c(c('Endpoint',apply(cbind(sq[-1], sq[-1]+11), 1, function(x) paste(x,collapse='-')))[-c(length(sq)-1,length(sq))], '>120'), include.lowest = T, ordered_result = T)

back.v$brks = with(back.v, factor(brks, rev(levels(brks))))

#back.v$Pathology = as.character(back.v$Pathology)
#back.v$Pathology = factor(back.v$Pathology, levels=c('BE','ID','LGD','HGD','IMC'), ordered = T)

@@

```{r echo=F, message=F, warning=F, fig.height=5, fig.width=8}
back.v = back.v %>% dplyr::mutate(Risk = factor(Risk, levels=c('High','Moderate','Low')), nl = factor(nl, ordered = T))

@@

@@

### 2019-08-14 Validation Report.Rmd
...

library(BarrettsProgressionRisk)
library(knitr)
library(kableExtra)
library(tidyverse)
library(gridExtra)
library(reshape2)

#source('lib/load_patient_metadata.R')

@@

@@

if (file.exists('batch8_preds.txt')) {
  batch8_preds = read_tsv('batch8_preds.txt')
} else {
  segFiles = grep(paste(on$`Patient ID`,collapse='|'), list.files('~/Data/BarrettsProgressionRisk/Analysis/validation', 'segObj',  full.names = T, recursive = T), value=T)

@@

    prr = BarrettsProgressionRisk::predictRiskFromSegments(segmented, F)
    BarrettsProgressionRisk::predictions(prr,'sample')
  }))
  batch8_preds = batch8_preds %>% filter(!is.na(`Patient ID`))

  write_tsv(batch8_preds, 'batch8_preds.txt')
}

@@

@@

if (file.exists('oldp_preds.txt')) {
  oldp.preds = read_tsv('oldp_preds.txt')
} else {
  segFiles = grep(paste(oldp$`Patient ID`,collapse='|'), list.files('~/Data/BarrettsProgressionRisk/Analysis/validation', 'segObj',  full.names = T, recursive = T), value=T)

@@

    prr = BarrettsProgressionRisk::predictRiskFromSegments(segmented,F)
    predictions(prr)
  }))
  write_tsv(oldp.preds, 'oldp_preds.txt')
}

@@

oldp.preds %>% dplyr::summarise(
  'TPR'=length(which(Risk == 'High'))/n(),
  'FNR'=length(which(Risk == 'Low'))/n()) %>% kable() %>% kable_styling(full_width = F)

oldp.preds %>% group_by(`Risk`) %>% dplyr::summarise( 'n samples'=n() ) %>% 
  kable(caption='Count sample predicted risks, P') %>% kable_styling(full_width = F)

fnr.patients = oldp.preds %>% 
  group_by(`Hospital Research ID`) %>% dplyr::summarise(
  TPR = length(which(Risk == 'High'))/n()
) %>% dplyr::summarise( 'Misclassified patients' = length(which(TPR < 0.5))/n() ) %>% pull

oldp.preds %>% group_by(`Hospital Research ID`) %>% dplyr::summarise(
  `n samples` = n(), 
  'High Risk' = length(which(Risk == 'High')),
  'Mod. Risk' = length(which(Risk == 'Moderate')),
  'Low Risk' =   length(which(Risk == 'Low')),
  'Pathology Notes' = paste(`Path Notes`, collapse='; ')
) %>% kable(caption = 'Classified predictions per patient (all P)') %>% kable_styling('striped')

@@

For the old progressor samples to-date, the model failed to classify any samples as high risk for *`r fnr.patients*100`%* of patients.  None of these were the samples I might have said were poor quality via manual QC so taken at face value these patients/samples were incorrectly predicted.

@@

if (file.exists('val_np.txt')) {
  new.np.preds = read_tsv('val_np.txt')
} else {
  segFiles = grep(paste(new.np$`Patient ID`,collapse='|'), list.files('~/Data/BarrettsProgressionRisk/Analysis/validation', '(2|3)_segObj',  full.names = T, recursive = T), value=T)

@@

    segmented$sample.info = BarrettsProgressionRisk::loadSampleInformation(info %>% filter(Sample %in% segmented$sample.info$Sample) )
    prr = BarrettsProgressionRisk::predictRiskFromSegments(segmented,F)
    predictions(prr)
  }))
  write_tsv(new.np.preds, 'val_np.txt')
}

@@

#cor.test(new.np.preds$n.segs, new.np.preds$varMAD_median)

new.np.preds %>% dplyr::summarise(
  'FPR'=length(which(Risk == 'High'))/n(),
  'TPR'=length(which(Risk == 'Low'))/n()) %>% kable() %>% kable_styling(full_width = F)

new.np.preds %>% group_by(`Risk`) %>% dplyr::summarise( 'n samples'=n() ) %>% 
  kable(caption='Count sample predicted risks, NP') %>% kable_styling(full_width = F)

fpr.patients = new.np.preds %>%  group_by(`Hospital Research ID`) %>% dplyr::summarise(
  TPR = length(which(Risk == 'Low'))/n()
) %>% dplyr::summarise( 'Misclassified patients' = length(which(TPR < 0.5))/n() ) %>% pull

@@

with(val.new.np.preds, cor.test(varMAD_median, Probability))

val.new.np.preds %>% group_by(`Hospital Research ID`) %>% dplyr::summarise(
  `n samples` = n(), 
  'High Risk' = length(which(Risk == 'High')),
  'Mod. Risk' = length(which(Risk == 'Moderate')),
  'Low Risk' =   length(which(Risk == 'Low')),
  'QC var' = signif( mean(varMAD_median), 2)
) %>% kable(caption = 'Classified predictions per patient (all NP)') %>% kable_styling('striped')

>>>>

library(tidyverse)
library(BarrettsProgressionRisk)
library(kableExtra)
library(gridExtra)
library(reshape2)

source('lib/common_plots.R')

@@

pred.dir = '~/Data/BarrettsProgressionRisk/Analysis/validation/predictions'
dir.create(pred.dir, showWarnings = F)

model = '~/Data/BarrettsProgressionRisk/Analysis/5e6_arms/all.pt.alpha.Rdata'
load(model, verbose=T)

fit = models$`0.9`
lambda = performance.at.1se$`0.9`$lambda

@@

if (file.exists(paste0(pred.dir, '/batch8_preds.txt'))) {
  batch8_preds = read_tsv(paste0(pred.dir, '/batch8_preds.txt'))
} else {
  segFiles = grep(paste(on$`Patient ID`,collapse='|'), list.files('~/Data/BarrettsProgressionRisk/Analysis/validation', 'segObj',  full.names = T, recursive = T), value=T)

@@

    prr = BarrettsProgressionRisk::predictRiskFromSegments(segmented, model = fit, s = lambda, verbose = F)
    BarrettsProgressionRisk::predictions(prr,'sample')
  }))
  batch8_preds = batch8_preds %>% filter(!is.na(`Patient ID`))

  write_tsv(batch8_preds, paste0(pred.dir, '/batch8_preds.txt'))
}

@@

The predictions from these samples are not particularly good.

@@

if (file.exists(paste0(pred.dir, '/oldp_preds.txt'))) {
  oldp.preds = read_tsv(paste0(pred.dir, '/oldp_preds.txt'))
} else {
  segFiles = grep(paste(oldp$`Patient ID`,collapse='|'), list.files('~/Data/BarrettsProgressionRisk/Analysis/validation', 'segObj',  full.names = T, recursive = T), value=T)

@@

    prr = BarrettsProgressionRisk::predictRiskFromSegments(segmented, model = fit, s = lambda, verbose = F)
    predictions(prr)
  }))
  write_tsv(oldp.preds, paste0(pred.dir, '/oldp_preds.txt'))
}

@@

oldp.preds %>% group_by(`Risk`) %>% dplyr::summarise( 'n samples'=n() ) %>% spread(Risk, `n samples`) %>%
  kable(caption='Progressor sample predicted risks') %>% kable_styling(full_width = F)

oldp.preds %>% group_by(`Hospital Research ID`) %>% dplyr::summarise(
  `n samples` = n(), 
  'High Risk' = length(which(Risk == 'High')),
  'Mod. Risk' = length(which(Risk == 'Moderate')),
  'Low Risk' =   length(which(Risk == 'Low')),
  'Pathology Notes' = paste(`Path Notes`, collapse='; ')
) %>% mutate(
  `High Risk` = cell_spec(`High Risk`, "html", color = ifelse(`High Risk` > 0, "red", "grey39")),
  `Mod. Risk` = cell_spec(`Mod. Risk`, "html", color = ifelse(`Mod. Risk` > 0, "orange", "grey39")),
  `Low Risk` = cell_spec(`Low Risk`, "html", color = ifelse(`Low Risk` > 0, "blue", "grey39")),
) %>% kable(format='html',escape = F,caption = 'Classified predictions per patient (all P)') %>% kable_styling('striped', full_width = F)

oldp.preds = oldp.preds %>% rowwise %>% mutate(PID = unlist(strsplit(`Block ID`, '(-|\\s)\\w+$', perl = T))[1])

ggplot(oldp.preds, aes(`Hospital Research ID`, Probability, color=Probability)) + geom_point() +
  scale_color_distiller(palette = 'RdYlBu', name='P(P)') + plot.theme + theme(axis.text.x = element_text(angle=45,hjust=1))

oldp.per.endo = oldp.preds %>% group_by(`Hospital Research ID`, PID) %>% 
  dplyr::mutate('n.samples'=length(PID),'Samples'=paste(Sample,collapse=',')) %>% 
  dplyr::summarise_at(vars(matches('Prob|Relative|Endo')), list(max)) %>% dplyr::select(-matches('^Sample$')) %>% 
  arrange(`Hospital Research ID`, Endoscopy, PID) 
#back.per.endo = back.per.endo %>% mutate(quants = cut(Prediction, breaks=cuts, include.lowest = T)) %>% left_join(pred.confidence %>% dplyr::select(quants, Risk), by='quants') %>% dplyr::select(-matches('quant')) %>% dplyr::mutate(Risk = as.character(Risk))

fnr.patients = oldp.preds %>% 
  group_by(`Hospital Research ID`) %>% dplyr::summarise(
  FNR = length(which(Risk == 'Low'))/n() ) %>% 
  dplyr::summarise( 'Misclassified patients' = length(which(FNR > 0.5))/n() ) %>% pull %>% round(2)

@@

For the old progressor samples to-date, the model failed to classify any samples as high or moderate risk for *`r fnr.patients*100`%* of patients.  None of these were the samples I might have said were poor quality via manual QC so taken at face value these patients/samples were incorrectly predicted.

@@

if (file.exists(paste0(pred.dir, '/val_np.txt'))) {
  new.np.preds = read_tsv(paste0(pred.dir, '/val_np.txt'))
} else {
  segFiles = grep(paste(sub('\\(.*\\)','', new.np$`Patient ID`),collapse='|'), list.files('~/Data/BarrettsProgressionRisk/Analysis/validation', '(2|3)_segObj',  full.names = T, recursive = T), value=T)

@@

    segmented$sample.info = BarrettsProgressionRisk::loadSampleInformation(info %>% filter(Sample %in% segmented$sample.info$Sample) )
    prr = BarrettsProgressionRisk::predictRiskFromSegments(segmented, model = fit, s = lambda, verbose = F)
    p = predictions(prr)
    print(p)
    return(p)
  }))
  write_tsv(new.np.preds, paste0(pred.dir, '/val_np.txt'))
}

@@

#cor.test(new.np.preds$n.segs, new.np.preds$varMAD_median)
#with(val.new.np.preds, cor.test(varMAD_median, Probability))

new.np.preds %>% group_by(`Risk`) %>% dplyr::summarise( 'n samples'=n() ) %>% spread(Risk, `n samples`) %>% 
  kable(caption='Count sample predicted risks, NP') %>% kable_styling(full_width = F)

@@

# val.new.np.preds %>% group_by(`Hospital Research ID`) %>% dplyr::summarise(
#   `n samples` = n(), 
#   'High Risk' =  length(which(Risk == 'High')),
#   'Mod. Risk' = length(which(Risk == 'Moderate')),
#   'Low Risk' =   length(which(Risk == 'Low')),
#   'QC var' = signif( mean(varMAD_median), 2)
# ) %>% mutate(
#   `High Risk` = cell_spec(`High Risk`, "html", color = ifelse(`High Risk` > 0, "red", "grey39")),
#   `Mod. Risk` = cell_spec(`Mod. Risk`, "html", color = ifelse(`Mod. Risk` > 0, "orange", "grey39")),
#   `Low Risk` = cell_spec(`Low Risk`, "html", color = ifelse(`Low Risk` > 0, "blue", "grey39")),
# ) %>%
#   kable(format='html',escape = F, caption = 'Classified predictions per patient (all NP)') %>% kable_styling('striped', full_width = F)

ggplot(new.np.preds, aes(Probability)) + geom_histogram(aes(fill=..x..), breaks=seq(0,1,0.1), show.legend = F) + scale_fill_distiller(palette = 'RdYlBu', name='P(P)') +
    plot.theme + labs(title='Sample predictions for validation NP samples', y='n Samples', x='Probability') + theme(legend.position = 'none')

new.np.preds = new.np.preds %>% rowwise %>% mutate(PID = unlist(strsplit(`Block ID`, '(-|\\s)\\w+$', perl = T))[1])

ggplot(new.np.preds, aes(`Hospital Research ID`, Probability, color=Probability)) + geom_point() +
  scale_color_distiller(palette = 'RdYlBu', name='P(P)') + plot.theme + theme(axis.text.x = element_text(angle=45,hjust=1))

np.per.endo = new.np.preds %>% group_by(`Hospital Research ID`, PID) %>% 
  dplyr::mutate('n.samples'=length(PID),'Samples'=paste(Sample,collapse=',')) %>% 
  dplyr::summarise_at(vars(matches('Prob|Relative|Endo')), list(max)) %>% dplyr::select(-matches('^Sample$')) %>% 
  arrange(`Hospital Research ID`, Endoscopy, PID) 
#back.per.endo = back.per.endo %>% mutate(quants = cut(Prediction, breaks=cuts, include.lowest = T)) %>% left_join(pred.confidence %>% dplyr::select(quants, Risk), by='quants') %>% dplyr::select(-matches('quant')) %>% dplyr::mutate(Risk = as.character(Risk))

#np.per.endo %>% filter(`Relative Risk` > 0)
